---
title: "Report Exercise Chapter 10"
output: html_document
date: "2023-05-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(rsample)
library(caret)
library(rsample)
library(recipes)
library(cowplot)


daily_fluxes_davos <- read_csv("FLX_CH-Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv")
daily_fluxes_lae <- read_csv("FLX_CH-Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv")

```

Loading of data and packages


```{r}
set.seed(123)  # for reproducibility

split <- rsample::initial_split(daily_fluxes_davos, prop = 0.8, strata = "VPD_F")
daily_fluxes_davos_train <- rsample::training(split)
daily_fluxes_davos_test <- rsample::testing(split)

split <- rsample::initial_split(daily_fluxes_lae, prop = 0.8, strata = "VPD_F")
daily_fluxes_lae_train <- rsample::training(split)
daily_fluxes_lae_test <- rsample::testing(split)



ggplot(davos_mod_cv)
ggplot(lae_mod_cv)
print(davos_mod_cv)
print(lae_mod_cv)

```
First model formulation, then train the model with the training data, using also the 10 fold cross validation and in the end also display the optimal k value.

```{r}

## metrics within site davos
davos_within_site <- eval_model(mod = davos_mod_cv, df_train = daily_fluxes_davos_train, df_test = daily_fluxes_davos_test)
print(davos_within_site)

# metrics within site laegern
laegern_within_site <- eval_model(mod = lae_mod_cv, df_train = daily_fluxes_lae_train, df_test = daily_fluxes_lae_test)
print(laegern_within_site)


#laegern acros site
laegern_across_site_eval <- eval_model(mod = davos_mod_cv,
                                       df_train = daily_fluxes_davos_train,
                                       df_test = daily_fluxes_lae_test)

print(laegern_across_site_eval)


#davos across site
davos_across_site_eval <- eval_model(mod = lae_mod_cv,
                                     df_train = daily_fluxes_lae_train,
                                     df_test = daily_fluxes_davos_test)
print(davos_across_site_eval)

